---
title: "Dendroclima"
author: "Muraja DOS"
date: "2023-04-07"
output: pdf_document
---

**Clear the console and the variables + tell your actual directory**
```{r}
cat("\014") 
rm(list=ls()) 
getwd() 
```

**Load libraries and change to the directory you want to work, which must have your data**
```{r}
library(dplR)
library(tidyverse)
library(ggplot2)
library(treeclim)
library(tidyverse)
library(hyfo)
library(ggpubr)
#setwd('C:/Users/DanyO/Documents/Dani_Documentos/Doutorado - Física e Astronomia_att10_01_2023/Softwares Programados/SFTW2_Dendroclimatology_Analysis')
```


**Monthly**
*Create the correlations + p-values but put it as a data.frame*
```{r}
tree_ring_df <- read.table('Chronology_Stabilized_Trees_canela.txt', header = TRUE)
SUNSPOT <- read.table("SUNSPOT_data.txt", header = FALSE)

Years <- SUNSPOT[,1]
row.names(SUNSPOT) <- Years
SUNSPOT <- SUNSPOT[,-1]
climatic_df <- SUNSPOT

# Set the significance level (alpha)
alpha <- 0.05

# Extract the relevant columns for correlation
rwi <- tree_ring_df[, 1]
climatic_data <- as.matrix(climatic_df)

# Extract the overlapping years between the two datasets
years <- intersect(rownames(tree_ring_df), rownames(climatic_df))

# Subset the tree ring chronology and climatic data based on the overlapping years
rwi <- rwi[match(years, rownames(tree_ring_df))]
climatic_data <- climatic_data[match(years, rownames(climatic_df)), ]

# Calculate correlations and significance for each month
results <- lapply(1:12, function(i) {
  month <- month.abb[i]
  correlation <- cor(rwi, climatic_data[, i], use = "complete.obs")
  p_value <- cor.test(rwi, climatic_data[, i], use = "complete.obs")$p.value
  significant <- p_value <= alpha
  
  # Remove the unwanted text from the correlation value
  correlation <- gsub(".*cor = ", "", as.character(correlation))
  
  data.frame(Month = month, Correlation = correlation, Pvalue = p_value, Significant = significant)
})

# Combine the results into a single data frame
results_df <- do.call(rbind, results)

# Display the results data frame
results_df

write.csv(results_df, "Correlation_SUNSPOTRWI.csv")
```

**SEASONAL**
*create the correlations + p-values but put it as a data.frame*
```{r}
tree_ring_df <- read.table('Chronology_Stabilized_trees_canela.txt', header = TRUE)
SUNSPOT <- read.table("SUNSPOT_data.txt", header = FALSE)

# Extract the relevant variables from the data
#rwi <- tree_ring_df$RWI[1:nrow(SUNSPOT)]  # Subset the first 31 rows of RWI

Years <- SUNSPOT[,1]
row.names(SUNSPOT) <- Years
SUNSPOT <- SUNSPOT[,-1]

# Define the months corresponding to each season
season_months <- list(
  Summer = c(12, 1, 2),
  Autumn = c(3, 4, 5),
  Winter = c(6, 7, 8),
  Spring = c(9, 10, 11)
)

# Create a function to calculate seasonal means
calculate_seasonal_means <- function(data, months) {
  seasons <- lapply(months, function(x) rowMeans(data[, x, drop = FALSE], na.rm = TRUE))
  seasons_df <- as.data.frame(do.call(cbind, seasons))
  colnames(seasons_df) <- names(months)
  return(seasons_df)
}

# Calculate seasonal means for temperature_med
SUNSPOT_seasons <- calculate_seasonal_means(SUNSPOT, season_months)

# Set the significance level (alpha)
alpha <- 0.05

# Extract the relevant columns for correlation
rwi <- tree_ring_df[, 1]
climatic_data <- as.matrix(SUNSPOT_seasons)

# Extract the overlapping years between the two datasets
years <- intersect(rownames(tree_ring_df), rownames(SUNSPOT_seasons))

# Subset the tree ring chronology and climatic data based on the overlapping years
rwi <- rwi[match(years, rownames(tree_ring_df))]
climatic_data <- climatic_data[match(years, rownames(SUNSPOT_seasons)), ]

# Calculate correlations and significance for each month
results <- lapply(1:4, function(i) {
  month <- month.abb[i]
  correlation <- cor(rwi, climatic_data[, i], use = "complete.obs")
  p_value <- cor.test(rwi, climatic_data[, i], use = "complete.obs")$p.value
  significant <- p_value <= alpha

  # Remove the unwanted text from the correlation value
  correlation <- gsub(".*cor = ", "", as.character(correlation))

  data.frame(Correlation = correlation, Pvalue = p_value)
})

# Combine the results into a single data frame
results_df <- do.call(rbind, results)

# Display the results data frame
results_df

write.csv(results_df, "CorrelationSUNSPOT_Seasonal.csv")
```


################################################################################


```{r}
cat("\014") 
rm(list=ls()) 
getwd() 
```

**SEMESTER**
*create the correlations + p-values but put it as a data.frame*
```{r}
tree_ring_df <- read.table('Chronology_Stabilized_trees_canela.txt', header = TRUE)
SUNSPOT <- read.table("SUNSPOT_data.txt", header = FALSE)

# Extract the relevant variables from the data
#rwi <- tree_ring_df$RWI[1:nrow(SUNSPOT)]  # Subset the first 31 rows of RWI

Years <- SUNSPOT[,1]
row.names(SUNSPOT) <- Years
SUNSPOT <- SUNSPOT[,-1]

# Define the merged seasons
semester_months <- list(
  Spring_Summer = c(9, 10, 11, 12, 1, 2),
  Autumn_Winter = c(3, 4, 5, 6, 7, 8)
)

# Create a function to calculate merged seasonal means
calculate_merged_semester_means <- function(data, months) {
  merged_semester <- lapply(months, function(x) rowMeans(data[, x, drop = FALSE], na.rm = TRUE))
  merged_semester_df <- as.data.frame(do.call(cbind, merged_semester))
  colnames(merged_semester_df) <- names(months)
  return(merged_semester_df)
}

# Calculate merged seasonal means for temperature_max
SUNSPOT_semester <- calculate_merged_semester_means(SUNSPOT, semester_months)


# Set the significance level (alpha)
alpha <- 0.05

# Extract the relevant columns for correlation
rwi <- tree_ring_df[, 1]
climatic_data <- as.matrix(SUNSPOT_semester)

# Extract the overlapping years between the two datasets
years <- intersect(rownames(tree_ring_df), rownames(SUNSPOT_semester))

# Subset the tree ring chronology and climatic data based on the overlapping years
rwi <- rwi[match(years, rownames(tree_ring_df))]
climatic_data <- climatic_data[match(years, rownames(SUNSPOT_semester)), ]

# Calculate correlations and significance for each month
results <- lapply(1:2, function(i) {
  month <- month.abb[i]
  correlation <- cor(rwi, climatic_data[, i], use = "complete.obs")
  p_value <- cor.test(rwi, climatic_data[, i], use = "complete.obs")$p.value
  significant <- p_value <= alpha

  # Remove the unwanted text from the correlation value
  correlation <- gsub(".*cor = ", "", as.character(correlation))

  data.frame(Correlation = correlation, Pvalue = p_value)
})

# Combine the results into a single data frame
results_df <- do.call(rbind, results)

# Display the results data frame
results_df

write.csv(results_df, "CorrelationSUNSPOT_Semester.csv")
```


################################################################################

```{r}
cat("\014") 
rm(list=ls()) 
getwd() 
```

**ANNUALY**
*create the correlations + p-values but put it as a data.frame*
```{r}
tree_ring_df <- read.table('Chronology_Stabilized_trees_canela.txt', header = TRUE)
SUNSPOT <- read.table("SUNSPOT_data.txt", header = FALSE)

# Extract the relevant variables from the data
#rwi <- tree_ring_df$RWI[1:nrow(SUNSPOT)]  # Subset the first 31 rows of RWI

Years <- SUNSPOT[,1]
row.names(SUNSPOT) <- Years
SUNSPOT <- SUNSPOT[,-1]


# Define the merged seasons
yearly_months <- list(
  Yearly = 1:12  # All months in a year
)

# Create a function to calculate merged seasonal means
calculate_merged_yearly_means <- function(data, months) {
  merged_yearly <- lapply(months, function(x) rowMeans(data[, x, drop = FALSE], na.rm = TRUE))
  merged_yearly_df <- as.data.frame(do.call(cbind, merged_yearly))
  colnames(merged_yearly_df) <- names(months)
  return(merged_yearly_df)
}

# Calculate merged seasonal means for temperature_max
SUNSPOT_yearly <- calculate_merged_yearly_means(SUNSPOT, yearly_months)

# Set the significance level (alpha)
alpha <- 0.05

# Your existing code...

# Extract the relevant columns for correlation
climatic_data <- SUNSPOT_yearly[, 1]
rwi <- tree_ring_df[, 1]

# Subset the climatic_data to match the length of rwi
climatic_data <- climatic_data[1:length(rwi)]

# Calculate correlations and significance for the temperature data
correlation <- cor(rwi, climatic_data, use = "complete.obs")
p_value <- cor.test(rwi, climatic_data, use = "complete.obs")$p.value
significant <- p_value <= alpha

# Remove the unwanted text from the correlation value
correlation <- gsub(".*cor = ", "", as.character(correlation))

# Create a data frame with the results
results_df <- data.frame(Correlation = correlation, Pvalue = p_value)

# Display the results data frame
print(results_df)

# Write the results to a CSV file
write.csv(results_df, "Correlation_SUNSPOT_Yearly.csv")

```

**Correlation Graphics**
#Remember to plot the graphics of the analysis you want. If you run the full code at once, the graphics will be only of your annual analysis. e.g. Run the first Chunk for monthly correlation and then come here to run the graphic Chunk.

*Bar Plot*
```{r}
# Convert 'Correlation' column to numeric
results_df$Correlation <- as.numeric(results_df$Correlation)

# Define color patterns based on months
color_patterns <- c("#f8766d", "#de8c00", "#b79f00", "#7cae00", "#00ba38", "#00c08b", 
                    "#00bfc4", "#00b4f0", "#619cff", "#c77cff", "#f564e3", "#ff64b0")

# Bar plot with colored bars and black line at correlation 0
barplot(results_df$Correlation, names.arg = results_df$Month,
        xlab = "Months", ylab = "Correlations",
        main = "Correlation between SUNSPOT and RWI - Canela",
        col = color_patterns)

# Add a black line at correlation 0
abline(h = 0, col = "black")


```

*Scatter Plot*
```{r}
# Convert 'Correlation' column to numeric
results_df$Correlation <- as.numeric(results_df$Correlation)

# Convert 'Month' column to factor with numeric levels
results_df$Month <- as.numeric(as.factor(results_df$Month))

# Define color patterns based on months
color_patterns <- c("#f8766d", "#de8c00", "#b79f00", "#7cae00", "#00ba38", "#00c08b", 
                    "#00bfc4", "#00b4f0", "#619cff", "#c77cff", "#f564e3", "#ff64b0")

# Create a scatter plot with switched axes
plot(results_df$Month, results_df$Correlation, type = "n",
     xlab = "Months", ylab = "Correlations",
     main = "Correlation between SUNSPOT and RWI - Canela",
     xaxt = "n", yaxt = "n")  # Disable default x and y axis labels

# Add scatter points
points(results_df$Month, results_df$Correlation, pch = 16, col = color_patterns)

# Add a black line at correlation 0
abline(h = 0, col = "black", lty = 2, lwd = 2)

# Add axis labels and tick marks manually
axis(1, at = 1:12, labels = month.abb, tick = TRUE, lwd.tick = 0)  # X-axis (months)
axis(2, las = 1, lwd.tick = 1)  # Y-axis

# Add a legend for the color patterns
legend("bottomright", legend = month.abb, fill = color_patterns, bty = "n", cex = 0.7)

# Add a title and adjust plot margins if needed
title(main = "Correlation between SUNSPOT and RWI - Canela", cex.main = 1.2)

```