---
title: "PartialCorrSeason.Rmd"
author: "Dani OS"
date: "2023-06-11"
output: pdf_document
---

#SEASONS WERE BASED ON http://estacao.iag.usp.br/seasons/index.php#:~:text=Ver%C3%A3o%3A%20Dezembro%2C%20Janeiro%20e%20Fevereiro%20(DJF)

**Clear the console and the variables + tell your actual directory**
```{r}
cat("\014") 
rm(list=ls()) 
getwd() 
```


**PARTIAL CORRELATION** 

**MONTHLY**

*PRECIPITATION EXCLUDING TMAX*
```{r}
# Load the required package
library(ppcor)

# Read the data
RWI <- read.table("Chronology_Stabilized_Trees_canela.txt", header = TRUE)
precipitation <- read.csv("PREC_CAXIAS_MEANIMP.csv")
temperature_max <- read.csv("TEMPMAX_CAXIAS_MEANIMP.csv")

# Extract the relevant variables from the data
rwi <- RWI$RWI[1:nrow(temperature_max)]  # Subset the first 31 rows of RWI
precip <- precipitation[1:nrow(temperature_max), -1]  # Subset the first 31 rows, exclude the Year column
tmax <- temperature_max[1:nrow(temperature_max), -1]  # Subset the first 31 rows, exclude the Year column

# Create a data frame to store correlation coefficients and p-values
cor_table <- data.frame(Month = colnames(precip), Correlation = numeric(ncol(precip)), p_value = numeric(ncol(precip)))

# Calculate the correlations and p-values for each month
for (i in 1:ncol(precip)) {
  data <- data.frame(RWI = rwi, Precipitation = precip[, i], TemperatureMax = tmax[, i])
  cor_result <- spcor(data, method = "pearson")
  cor_table[i, "Correlation"] <- cor_result$estimate[1, 2]
  cor_table[i, "p_value"] <- cor_result$p.value[1, 2]
}

# Print the correlation table
print(cor_table)
write.csv(cor_table,"MONTHLYpartialcorr_PRCPTMAX.csv")
cor_table_monthly_prcptmax <- cor_table
```


*PRECIPITATION EXCLUDING TMED*
```{r}
# Load the required package
library(ppcor)

# Read the data
RWI <- read.table("Chronology_Stabilized_Trees_canela.txt", header = TRUE)
precipitation <- read.csv("PREC_CAXIAS_MEANIMP.csv")
temperature_mean <- read.csv("TEMPMED_CAXIAS_MEANIMP.csv")

# Extract the relevant variables from the data
rwi <- RWI$RWI[1:nrow(temperature_max)]  # Subset the first 31 rows of RWI
precip <- precipitation[1:nrow(temperature_max), -1]  # Subset the first 31 rows, exclude the Year column
tmean <- temperature_mean[1:nrow(temperature_max), -1]  # Subset the first 31 rows, exclude the Year column

# Create a data frame to store correlation coefficients and p-values
cor_table <- data.frame(Month = colnames(precip), Correlation = numeric(ncol(precip)), p_value = numeric(ncol(precip)))

# Calculate the correlations and p-values for each month
for (i in 1:ncol(precip)) {
  data <- data.frame(RWI = rwi, Precipitation = precip[, i], TemperatureMean = tmean[, i])
  cor_result <- spcor(data, method = "pearson")
  cor_table[i, "Correlation"] <- cor_result$estimate[1, 2]
  cor_table[i, "p_value"] <- cor_result$p.value[1, 2]
}

# Print the correlation table
print(cor_table)
write.csv(cor_table, "MONTHLYpartialcorr_PRCPTMEAN.csv")
cor_table_monthly_prcptmed <- cor_table
```

*PRECIPITATION EXCLUDING TMIN*
```{r}
# Load the required package
library(ppcor)

# Read the data
RWI <- read.table("Chronology_Stabilized_Trees_canela.txt", header = TRUE)
precipitation <- read.csv("PREC_CAXIAS_MEANIMP.csv")
temperature_min <- read.csv("TEMPMIN_CAXIAS_MEANIMP.csv")

# Extract the relevant variables from the data
rwi <- RWI$RWI[1:nrow(temperature_max)]  # Subset the first 31 rows of RWI
precip <- precipitation[1:nrow(temperature_max), -1]  # Subset the first 31 rows, exclude the Year column
tmin <- temperature_min[1:nrow(temperature_max), -1]  # Subset the first 31 rows, exclude the Year column

# Create a data frame to store correlation coefficients and p-values
cor_table <- data.frame(Month = colnames(precip), Correlation = numeric(ncol(precip)), p_value = numeric(ncol(precip)))

# Calculate the correlations and p-values for each month
for (i in 1:ncol(precip)) {
  data <- data.frame(RWI = rwi, Precipitation = precip[, i], TemperatureMin = tmin[, i])
  cor_result <- spcor(data, method = "pearson")
  cor_table[i, "Correlation"] <- cor_result$estimate[1, 2]
  cor_table[i, "p_value"] <- cor_result$p.value[1, 2]
}

# Print the correlation table
print(cor_table)
write.csv(cor_table, "MONTHLYpartialcorr_PRCPTMIN.csv")
cor_table_monthly_prcptmin <- cor_table
```


**Graphic for Monthly partial correlation**
```{r}
library(ggplot2)
library(cowplot)

# Function to generate correlation plots
generate_correlation_plot <- function(correlation_table, title) {
  plot <- ggplot(correlation_table, aes(x = Month, y = Correlation, fill = Month)) +
    geom_bar(stat = "identity", position = "dodge", color = "black") +
    labs(title = title, x = "Months", y = "Correlation") +
    theme_minimal() +
    theme(legend.position = "none")
  
  return(plot)
}

# Create correlation plots for each table
plot_monthly_prcptmax <- generate_correlation_plot(cor_table_monthly_prcptmax, "Canela Monthly Correlation RWI - Prec. excluding TMax")
plot_monthly_prcptmed <- generate_correlation_plot(cor_table_monthly_prcptmed, "Canela Monthly Correlation RWI - Prec. excluding TMean")
plot_monthly_prcptmin <- generate_correlation_plot(cor_table_monthly_prcptmin, "Canela Monthly Correlation RWI - Prec. excluding TMin")

# Combine the plots into a single figure with rows
figure <- cowplot::plot_grid(
  plot_monthly_prcptmax, 
  plot_monthly_prcptmed, 
  plot_monthly_prcptmin,
  ncol = 1, labels = c("A", "B", "C")
)

# Display the combined plot
print(figure)

```


################################################################################
```{r}
cat("\014") 
rm(list=ls()) 
getwd() 
```

**SEASONAL**

*PARTIAL CORRELATION WITH PRECIPITATION EXCLUDING TMAX*
```{r}
# Load the required package
library(ppcor)

# Read the data
RWI <- read.table("Chronology_Stabilized_Trees_canela.txt", header = TRUE)
precipitation <- read.csv("PREC_CAXIAS_MEANIMP.csv")
temperature_max <- read.csv("TEMPMAX_CAXIAS_MEANIMP.csv")

# Extract the relevant variables from the data
rwi <- RWI$RWI[1:nrow(temperature_max)]  # Subset the first 31 rows of RWI

# Define the months corresponding to each season
season_months <- list(
  Summer = c(12, 1, 2),
  Autumn = c(3, 4, 5),
  Winter = c(6, 7, 8),
  Spring = c(9, 10, 11)
)

# Create a function to calculate seasonal means
calculate_seasonal_means <- function(data, months) {
  seasons <- lapply(months, function(x) rowMeans(data[, x, drop = FALSE], na.rm = TRUE))
  seasons_df <- as.data.frame(do.call(cbind, seasons))
  colnames(seasons_df) <- names(months)
  return(seasons_df)
}

# Calculate seasonal means for precipitation
precip_seasons <- calculate_seasonal_means(precipitation[1:nrow(temperature_max), -1], season_months)

# Calculate seasonal means for temperature_max
tmax_seasons <- calculate_seasonal_means(temperature_max[1:nrow(temperature_max), -1], season_months)

# Create a data frame to store correlation coefficients and p-values
cor_table <- data.frame(Month = colnames(precip_seasons), Correlation = numeric(ncol(precip_seasons)), p_value = numeric(ncol(precip_seasons)))

# Calculate the correlations and p-values for each month
for (i in 1:ncol(precip_seasons)) {
  data <- data.frame(RWI = rwi, Precipitation = precip_seasons[, i], TemperatureMax = tmax_seasons[, i])
  cor_result <- spcor(data, method = "pearson")
  cor_table[i, "Correlation"] <- cor_result$estimate[1, 2]
  cor_table[i, "p_value"] <- cor_result$p.value[1, 2]
}

# Print the correlation table
print(cor_table)
write.csv(cor_table,"SEASONALpartialcorr_PRCPTMAX.csv")
cor_table_seasonal_prcptmax <- cor_table
```

*PARTIAL CORRELATION WITH PRECIPITATION EXCLUDING TMED*
```{r}
# Load the required package
library(ppcor)

# Read the data
RWI <- read.table("Chronology_Stabilized_Trees_canela.txt", header = TRUE)
precipitation <- read.csv("PREC_CAXIAS_MEANIMP.csv")
temperature_mean <- read.csv("TEMPMED_CAXIAS_MEANIMP.csv")

# Extract the relevant variables from the data
rwi <- RWI$RWI[1:nrow(temperature_max)]  # Subset the first 31 rows of RWI

# Define the months corresponding to each season
season_months <- list(
  Summer = c(12, 1, 2),
  Autumn = c(3, 4, 5),
  Winter = c(6, 7, 8),
  Spring = c(9, 10, 11)
)

# Create a function to calculate seasonal means
calculate_seasonal_means <- function(data, months) {
  seasons <- lapply(months, function(x) rowMeans(data[, x, drop = FALSE], na.rm = TRUE))
  seasons_df <- as.data.frame(do.call(cbind, seasons))
  colnames(seasons_df) <- names(months)
  return(seasons_df)
}

# Calculate seasonal means for precipitation
precip_seasons <- calculate_seasonal_means(precipitation[1:nrow(temperature_max), -1], season_months)

# Calculate seasonal means for temperature_max
tmean_seasons <- calculate_seasonal_means(temperature_mean[1:nrow(temperature_max), -1], season_months)

# Create a data frame to store correlation coefficients and p-values
cor_table <- data.frame(Month = colnames(precip_seasons), Correlation = numeric(ncol(precip_seasons)), p_value = numeric(ncol(precip_seasons)))

# Calculate the correlations and p-values for each month
for (i in 1:ncol(precip_seasons)) {
  data <- data.frame(RWI = rwi, Precipitation = precip_seasons[, i], TemperatureMean = tmean_seasons[, i])
  cor_result <- spcor(data, method = "pearson")
  cor_table[i, "Correlation"] <- cor_result$estimate[1, 2]
  cor_table[i, "p_value"] <- cor_result$p.value[1, 2]
}

# Print the correlation table
print(cor_table)
write.csv(cor_table,"SEASONALpartialcorr_PRCPTMEAN.csv")
cor_table_seasonal_prcptmean <- cor_table

```


*PARTIAL CORRELATION WITH PRECIPITATION EXCLUDING TMIN*
```{r}
# Load the required package
library(ppcor)

# Read the data
RWI <- read.table("Chronology_Stabilized_Trees_canela.txt", header = TRUE)
precipitation <- read.csv("PREC_CAXIAS_MEANIMP.csv")
temperature_min <- read.csv("TEMPMIN_CAXIAS_MEANIMP.csv")

# Extract the relevant variables from the data
rwi <- RWI$RWI[1:nrow(temperature_max)]  # Subset the first 31 rows of RWI

# Define the months corresponding to each season
season_months <- list(
  Summer = c(12, 1, 2),
  Autumn = c(3, 4, 5),
  Winter = c(6, 7, 8),
  Spring = c(9, 10, 11)
)

# Create a function to calculate seasonal means
calculate_seasonal_means <- function(data, months) {
  seasons <- lapply(months, function(x) rowMeans(data[, x, drop = FALSE], na.rm = TRUE))
  seasons_df <- as.data.frame(do.call(cbind, seasons))
  colnames(seasons_df) <- names(months)
  return(seasons_df)
}

# Calculate seasonal means for precipitation
precip_seasons <- calculate_seasonal_means(precipitation[1:nrow(temperature_max), -1], season_months)

# Calculate seasonal means for temperature_min
tmin_seasons <- calculate_seasonal_means(temperature_min[1:nrow(temperature_max), -1], season_months)

# Create a data frame to store correlation coefficients and p-values
cor_table <- data.frame(Month = colnames(precip_seasons), Correlation = numeric(ncol(precip_seasons)), p_value = numeric(ncol(precip_seasons)))

# Calculate the correlations and p-values for each month
for (i in 1:ncol(precip_seasons)) {
  data <- data.frame(RWI = rwi, Precipitation = precip_seasons[, i], TemperatureMax = tmin_seasons[, i])
  cor_result <- spcor(data, method = "pearson")
  cor_table[i, "Correlation"] <- cor_result$estimate[1, 2]
  cor_table[i, "p_value"] <- cor_result$p.value[1, 2]
}

# Print the correlation table
print(cor_table)
write.csv(cor_table,"SEASONALpartialcorr_PRCPTMIN.csv")
cor_table_seasonal_prcptmin <- cor_table
```

**Graphic for Seasonal partial correlation**
```{r}
library(ggplot2)
library(cowplot)

# Function to generate correlation plots
generate_correlation_plot <- function(correlation_table, title) {
  plot <- ggplot(correlation_table, aes(x = Month, y = Correlation, fill = Month)) +
    geom_bar(stat = "identity", position = "dodge", color = "black") +
    labs(title = title, x = "Months", y = "Correlation") +
    theme_minimal() +
    theme(legend.position = "none")
  
  return(plot)
}

# Create correlation plots for each table
plot_monthly_prcptmax <- generate_correlation_plot(cor_table_seasonal_prcptmax, "Canela Seasonal Correlation RWI - Prec. excluding TMax")
plot_monthly_prcptmed <- generate_correlation_plot(cor_table_seasonal_prcptmean, "Canela Seasonal Correlation RWI - Prec. excluding TMean")
plot_monthly_prcptmin <- generate_correlation_plot(cor_table_seasonal_prcptmin, "Canela Seasonal Correlation RWI - Prec. excluding TMin")

# Combine the plots into a single figure with rows
figure <- cowplot::plot_grid(
  plot_monthly_prcptmax, 
  plot_monthly_prcptmed, 
  plot_monthly_prcptmin,
  ncol = 1, labels = c("A", "B", "C")
)

# Display the combined plot
print(figure)
```
